Parameter estimation

### 2023-04-27 SFE-article/Article_Model_Parameters.tex
27.04

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
We consider a system of partial differential equations, $G(t,x,\dot{x},\Theta,u)=0$, where $x$ denotes the state variables, $\Theta$ represents the parameters, and $u$ is the control variables. Our focus is on the supercritical extraction process, which involves a partially filled extractor with a fixed bed that operates under constant conditions. We assume the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, we use the concept of quasi-one-dimensional flow.
To describe the fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent, we use a distributed-parameter model based on \citet{Reverchon1996}. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and saturation concentration. We estimate these parameters from four experiments at different temperatures and pressures: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. We obtain the parameter values using the maximum likelihood estimation method on yield data under the normal error assumption.

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
We consider a system of partial differential equations, $G(t,x,\dot{x},\Theta,u)=0$, where $x$ denotes the state variables, $\Theta$ represents the parameters, and $u$ is the control variables. Our focus is on the supercritical extraction process, which involves a partially filled extractor with a fixed bed that operates under constant conditions. We assume the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, we use the concept of quasi-one-dimensional flow.
To describe the fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent, we use a distributed-parameter model based on \citet{Reverchon1996}. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and an initial state estimate. We estimate these parameters by applying the maximum likelihood estimation method on yield data under the normal error assumption.
The dataset comes from four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar.

### 2023-04-27 SFE-article/Sections/Results.tex
27.04

@@

    Moreover, the initial state estimation was performed together with parameter estimation. The concentration of solute in the solid phase is assumed to be constant and uniformly distributed. On the other hand, the solute concentration in the fluid phase is assumed not to be uniform and not zero. During the preparation period, when the operating conditions are achieved, the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed) to the region where there is no solid phase. As a result, the distribution of solutes mass in the fluid phase is assumed to not be uniform, and described by an arbitrary function $H=H(z)$. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in appendix \ref{CH: IC_BC}. The linear distribution can be defined if the total mass of solute $m_{total}$ and partition coefficient $\tau$ are know.

@@

	The parameter estimation was run as constrained optimization problem with following constraints:

	\begin{itemize}
		\item $0  \leq k_m   \leq 1$
		\item $0  \leq D_i   \leq +\infty$
		\item $0  \leq D_e^M \leq +\infty$
		\item $76 \leq m_{total} \leq 100$
		\item $0  \leq \tau \leq 1$
		\item $0  \leq \sigma \leq +\infty$
	\end{itemize}

@@

	\begin{table}[H]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautotabular{Figures/Results_estimation/estimation.csv} }
		\caption{Parameter estimation results}
		\label{tab: Estimation_results}
	\end{table}

>>>>

	To ensure the stability of the convergence while solving the partial differential equations, the Courant–Friedrichs–Lewy condition (given by equation) was applied. Based on the discretization step and the velocity, the Courant–Friedrichs–Lewy condition can be used to find.

	{\footnotesize
		\begin{equation*}
			C = \cfrac{u\Delta t}{\Delta x} \leq C_{max} = 1
		\end{equation*}
	}

@@

    Moreover, the initial state estimation was performed together with parameter estimation. The concentration of solute in the solid phase is assumed to be constant and uniformly distributed. On the other hand, the solute concentration in the fluid phase should not follow the same assumption as the solid phase. During the preparation period the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed) along the extractor. As a result, the distribution of solutes mass in the fluid phase is assumed to not be uniform. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in appendix \ref{CH: IC_BC}. The linear distribution can be defined if the total mass of solute $m_{total}$ and partition coefficient $\tau$ are know.

@@

	To ensure that parameters found by the optimizer do not reach unrealistic values, an additional set of inequality constraints is introduced. The lower and upper bounds for each parameter are given in Table \ref{tab:Constraints}.

	\begin{table}[!h]
		\begin{tabular}{ |l|c|c|c|c|c|c| } 
			\hline
			Parameter	&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Lower bound	&0	  	& 0 	  	& 0 		& 76 		 	& 0 	   	& 0 \\  \hline
			Upper bound	&1 		& $+\infty$ & $+\infty$	& 100 			& 1 		& $+\infty$ \\
			\hline
		\end{tabular}
	\caption{Constraints of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

	The parameter estimation problem requires an initial guess as an input to an optimization problem. To ensure that the solution is a global solution multiple initial points have been tested. The initial guesses for the parameter estimation problem are given in Table \ref{tab:InitialGuess}

	\begin{table}[!h]
		\begin{tabular}{ |l|c|c|c|c|c|c| } 
			\hline
			Parameter		&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Initial guess	&1	  	& 3 	  	& 1 		& 77 		 	& 0.65 	   	& 0.1 \\  
			\hline
		\end{tabular}
		\caption{Constraints of the parameter estimation problem}
		\label{tab:InitialGuess}
	\end{table}

@@

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautotabular{Figures/Results_estimation/estimation.csv} }
		\caption{Parameter estimation results}
		\label{tab: Estimation_results}
	\end{table}

### 2023-04-27 SFE-article/Sections/Thermodynamic_details.tex
27.04

	\subsubsection{Equation of state and properties of the fluid phase} \label{subsubsec: Equation of state}
	We consider equations of state in the general form ${\color{red}P}(t) {\color{orange}v_m}\left[T(t,z),P(t)\right] = {\color{orange}Z}{\color{magenta}R}{\color{blue}T}(t,z)$, where ${\color{magenta}v_m}$ denotes the molar volume of $\text{CO}_2$, ${\color{orange}Z}$ represents its compressibility factor, and $R$ is the universal gas constant. More specifically, we are interested in these equations because of the possibility to express the compressibility $Z$ as an explicit function of temperature and pressure. This is the case when $Z$ is obtained as one of the physically meaningful roots of a polynomial equation, like the Peng-Robinson's and {\color{red} stuff with REFS to equations of state}. \\

	In Peng-Robinson's equation, the compressibility ${\color{orange}\overline{Z}}\left[{\color{blue}T}(t,z), {\color{red}P}(t)\right]$ solves the third-order polynomial equation \todo{reference to Cardano formula}

@@

	\subsubsection{Departure functions for enthalpy calculations} \label{CH:Enthalpy} \todo[]{Change to my text}

	In thermodynamics, a departure function is defined for any thermodynamic property as the difference between the property as computed for an ideal gas and the property of the species as it exists in the real world, for a specified temperature T and pressure P. Common departure functions include those for enthalpy, entropy, and internal energy.

	Departure functions are used to calculate real fluid extensive properties (i.e. properties which are computed as a difference between two states). A departure function gives the difference between the real state, at a finite volume or non-zero pressure and temperature, and the ideal state, usually at zero pressure or infinite volume and temperature.

	For example, to evaluate enthalpy change between two points $h(V_1,T_1)$ and $h(V_2,T_2)$ we first compute the enthalpy departure function between volume $V_1$ and infinite volume at $T = T_1$, then add to that the ideal gas enthalpy change due to the temperature change from $T_1$ to $T_2$, then subtract the departure function value between $V_2$ and infinite volume.

	Departure functions are computed by integrating a function which depends on an equation of state and its derivative. The general form the enthalpy equation is given by

@@

>>>>

	\subsubsection{Equation of state and properties of the fluid phase} \label{subsubsec: Equation of state}
	We consider equations of state in the general form ${\color{red}P}(t) {\color{orange}v_m}[T(t,z),P(t)] = {\color{orange}Z}{\color{magenta}R}{\color{blue}T}(t,z)$, where ${\color{magenta}v_m}$ denotes the molar volume of $\text{CO}_2$, ${\color{orange}Z}$ represents its compressibility factor, and $R$ is the universal gas constant. More specifically, we are interested in these equations because of the possibility to express the compressibility $Z$ as an explicit function of temperature and pressure. This is the case when $Z$ is obtained as one of the physically meaningful roots of a polynomial equation, like the Van der Waals's, Peng-Robinson's or Redlich–Kwong's equation of state. In this work the Peng-Robinson equation of state (\citet{Peng1976}) is used

	In Peng-Robinson's equation fo state, the compressibility ${\color{orange}\overline{Z}}[{\color{blue}T}(t,z), {\color{red}P}(t)]$ solves the third-order polynomial equation given by equitation \ref{eq: Compressibility}. The polynomial equation can be solved by numerical methods like Newton-Raphson, or by applying Cardano formula. The Cardano formula is preferable in comparison to numerical root-finders, as it allows to obtain an analytical solution. The details of the Cardano formula are presented in appendix \ref{CH: Cardano}.

@@

	\subsubsection{Departure functions for enthalpy calculations} \label{CH:Enthalpy}

	In thermodynamics, a departure function is a concept used to calculate the difference between a real fluid's thermodynamic properties and those of an ideal gas, given a specific temperature and pressure. Common departure functions include those for enthalpy, entropy, and internal energy. These functions are used to calculate extensive properties, which are properties computed as a difference between two states.

	For example, to evaluate the enthalpy change between two points, $h(V_1,T_1)$ and $h(V_2,T_2)$, we first calculate the enthalpy departure function between the initial volume $V_1$ and infinite volume at temperature $T_1$. We then add to that the ideal gas enthalpy change due to the temperature change from $T_1$ to $T_2$, and finally subtract the departure function value between the final volume $V_2$ and infinite volume.

	Departure functions are computed by integrating a function that depends on an equation of state and its derivative. The general form of the enthalpy equation is given by:

@@

	Here, $h^{id}$ represents the enthalpy of an ideal gas, $h$ is the enthalpy of a real fluid, $R$ is the universal gas constant, $T$ is temperature, $v_m$ is the molar volume, and $Z$ is the compressibility factor.

	The integral in the equation is evaluated over the range of molar volumes from $v_m$ to infinity. The integral includes a term that depends on the derivative of the compressibility factor with respect to temperature, evaluated at the molar volume $v_m$. Finally, the term $1-Z$ is added to account for the deviation of the fluid's properties from those of an ideal gas.
